\pdfoutput=1
%% ****** Start of file apstemplate.tex ****** %
%%
%%
%% This file is part of the APS files in the REVTeX 4 distribution.
%% Version 4.1r of REVTeX, August 2010
%%
%%
%% Copyright (c) 2001, 2009, 2010 The American Physical Society.
%%
%% See the REVTeX 4 README file for restrictions and more information.
%%
%
% This is a template for producing manuscripts for use with REVTEX 4.0
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.
%
% Group addresses by affiliation; use superscriptaddress for long
% author lists, or if there are many overlapping affiliations.
% For Phys. Rev. appearance, change preprint to twocolumn.
% Choose pra, prb, prc, prd, pre, prl, prstab, prstper, or rmp for journal
% Add 'draft' option to mark overfull boxes with black boxes
% Add 'showpacs' option to make PACS codes appear
% Add 'showkeys' option to make keywords appear
\documentclass[aps,prd,twocolumn,superscriptaddress,preprintnumbers,floatfix,nofootinbib]{revtex4-2}
%\documentclass[aps,prl,preprint,superscriptaddress]{revtex4-1}
%\documentclass[aps,prl,reprint,groupedaddress]{revtex4-1}

\usepackage{graphicx}
\usepackage{amsmath}
%\usepackage{mdwlist}
\usepackage[caption=false]{subfig}
\usepackage{siunitx}
\usepackage{placeins}
\usepackage{color}
\usepackage{standalone}
\usepackage{dcolumn}
\usepackage{tensor}
\usepackage{bm}
%\usepackage{MnSymbol}
\usepackage{microtype}
\usepackage{etoolbox}
\usepackage{amssymb}
\usepackage{mathrsfs}
\usepackage{accents}
\usepackage[normalem]{ulem}
\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks,urlcolor=NavyBlue,citecolor=NavyBlue,linkcolor=NavyBlue,pdfusetitle]{hyperref}
\usepackage[all]{hypcap}
\usepackage[inline]{enumitem}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage{array}
\usepackage{booktabs}

\newcommand{\ts}{\textsuperscript}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\newcommand{\nn}{\nonumber}

\newcommand*{\eq}[1]{Eq.~\eqref{eq:#1}}
\newcommand*{\fig}[1]{Fig.~\ref{fig:#1}}
\newcommand*{\sect}[1]{Sec.~\ref{sec:#1}}

\newcommand{\boldmu}{\boldsymbol{\mu}}
\newcommand{\boldn}{\mathbf{n}}
\newcommand{\boldd}{\mathbf{d}}
\newcommand{\bolds}{\mathbf{s}}
\newcommand{\boldSigma}{\boldsymbol{\Sigma}}

\newcommand{\area}{\mathcal{A}}

%% Try to control orphans, widows, and extra whitespace
%\widowpenalty=10000
%\clubpenalty=10000
%\raggedbottom
\interfootnotelinepenalty=3000

%% Enable/disable comments
\newtoggle{commentsoff}
\togglefalse{commentsoff}
%\toggletrue{commentsoff}

% Check for command-line option to turn comments off
% https://tex.stackexchange.com/questions/1492/passing-parameters-to-a-document
\ifdefined\nocomments
    \toggletrue{commentsoff}
\fi

\iftoggle{commentsoff}{
  \newcommand*{\mi}[1]{}
  \newcommand*{\mg}[1]{}
  \newcommand*{\wf}[1]{}
  \newcommand*{\comment}[1]{}
  \newcommand*{\suggest}[3]{#2}
  \newcommand*{\todo}[1]{}
  \newcommand*{\warn}[1]{}
  \newcommand*{\red}[1]{#1}
  \newcommand*{\blue}[1]{#1}
  \newcommand{\cn}{}
  \newcommand*{\commentmark}[1]{}
}{
  \newcommand*{\mi}[1]{{\color{magenta} [{\bf MAX}: #1]}}
  \newcommand*{\wf}[1]{\textcolor{green}{\textbf{WILL}: #1}}
  \newcommand*{\suggest}[3]{\textcolor{Purple}{\textbf{#3} \sout{#1} #2}}
  \newcommand*{\comment}[1]{{\color{blue} [{\bf NOTE}: #1]}}
  \newcommand*{\warn}[1]{{\color{red} [{\bf WARNING}: #1]}}
  \newcommand*{\todo}[1]{{\color{red} [{\bf TODO}: #1]}}
  \newcommand*{\red}[1]{{\color{purple} #1}}
  \newcommand*{\blue}[1]{{\color{blue} #1}}
  \newcommand{\cn}{\blue{\bf [cn]}}
}

\graphicspath{{./fig/}}

\newcommand{\dcc}{LIGO-PXXXXXXX}

% NOTATION MACROS
\newcommand{\infd}{\mathrm{d}}
\newcommand{\white}{\bar}

\newcommand{\snropt}{\mathrm{SNR}}
\newcommand{\snrmf}{\mathrm{SNR}_{\rm mf}}
\newcommand{\cov}{C}
\newcommand{\acf}{\rho}
\newcommand{\nmode}{D}

\begin{document}

% Use the \preprint command to place your local institutional report
% number in the upper righthand corner of the title page in preprint mode.
% Multiple \preprint commands are allowed.
% Use the 'preprintnumbers' class option to override journal defaults
% to display numbers if necessary
% \preprint{\dcc}

%Title of paper
\title{Parametrizing gravitational-wave polarizations}

\author{Maximiliano Isi}
\email[]{maxisi@mit.edu}
\thanks{NHFP Einstein fellow}
%\homepage[]{Your web page}
%\thanks{}
%\altaffiliation{}
\affiliation{
LIGO Laboratory, Massachusetts Institute of Technology, Cambridge, Massachusetts 02139, USA
}%
%\affiliation{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010}

% Because hyperref only gets the *last* author, we need to be explicit.
\hypersetup{pdfauthor={Isi}}

\date{\today}

\begin{abstract}
Amazing!
\end{abstract}

\maketitle

% \tableofcontents

\section{Introduction}
\label{sec:intro}

It is useful to parametrize gravitational-wave (GW) polarizations in different ways.
Here I outline some useful alternatives and show how they are related.
It is important to understand the priors implied by the choice of parametrization through the corresponding Jacobians.

\section{General relativity}

In GR, there exist two propagating gravitational degrees of freedom, corresponding to two independent GW polarizations.
Their local effect can be encoded in a driving matrix $h_{ij}$ corresponding to the transverse-traceless part of the metric perturbation.
In a frame with $z$-axis along the direction of propagation, we can write this as
\beq \label{eq:hij}
(h_{ij}) = \begin{pmatrix}
h_+ & h_\times  & 0 \\
h_\times  & - h_+ & 0  \\
0 & 0 & 0
\end{pmatrix} ,
\eeq
where the plus ($h_+$) and cross ($h_\times$) polarizations are implicit functions of the retarded time $t - R/c$ to be specified by the source dynamics and the distance $R$ to the source.

Equation \eqref{eq:hij} presumes a specific choice of frame orientation.
As mentioned above, $h_{ij}$ only takes this simple form if our $z$ axis is parallel to the (spatial) wave vector $\vec{k}$.
However, there is no restriction regarding the orientation of the $x$ and $y$ axes within the plane perpendicular to $\vec{k}$; this is usually specified by an arbitrary \emph{polarization angle} $\psi$, defined with respect to some convenient reference (e.g., the celestial equator).
Although a choice of $\psi$ is required to prescribe $h_+$ and $h_\times$ in \eq{hij}, the form of this equation is invariant under changes in $\psi$.
To see this clearly, re-express \eq{hij} as 
\beq
h_{ij} = h_+ e^+_{ij} + h_\times e^\times_{ij}\,\,
\eeq
and construct the polarization tensors $e^{+/\times}_{ij}$ from two orthonormal vectors $\{\hat{x},\hat{y}\}$ perpendicular to $\hat{z}$,
\beq
e^+_{ij} \equiv \hat{x}_i \hat{x}_j - \hat{x}_i \hat{x}_j \, ,
\eeq
\beq
e^\times_{ij} \equiv \hat{x}_i \hat{y}_j + \hat{x}_i \hat{y}_j\, .
\eeq
By considering the behavior of $\hat{x}$ and $\hat{y}$ under rotations around $\hat{z}$, it is straightforward to show that the effect of rotating the polarization frame by some angle $\Delta\Psi$ amounts to redefining
\beq \label{eq:htransf}
h_+ \rightarrow h_+' = h_+ \cos 2\Delta \psi - h_\times \sin 2\Delta\psi \, ,
\eeq
\beq
h_\times \rightarrow h_\times' = h_\times \cos 2\Delta \psi + h_+ \sin 2\Delta\psi \, ,
\eeq
in \eq{hij}.
Therefore, $h_+$ and $h_\times$ are nothing but the two components of a tensor field with spin-weight $s=2$, and the two polarizations are only defined up to an arbitrary choice of $\psi$.

In the small-antenna limit, the signal induced by a passing GW on a given detector can be written as the projection
\beq
h(t) \equiv D^{ij} h_{ij} = F_+ h_+ + F_\times h_\times\, 
\eeq
with antenna patterns $F_{+/\times} \equiv D^{ij} e^{+/\times}_{ij}$ defined in terms of a detector tensor $D_{ij}$ representing the geometry of the measurement.
For a differential-arm detector with arms pointing along unit vectors $\hat{X}$ and $\hat{Y}$, this is just $D_{ij} = (\hat{X}_i \hat{X}_j - \hat{Y}_i \hat{Y}_j)/2$.
The antenna patterns are thus purely geometric factors that encode the relative orientations of the detector and wave frames, as defined by $\{\hat{X},\, \hat{Y},\, \hat{Z}\}$ and $\{\hat{x},\, \hat{y},\, \hat{z}\}$ respectively.
Under rotations of the wave frame, they transform through an expression complementary to \eq{htransf},
\beq \label{eq:Ftransf}
h_+ \rightarrow F_+' = F_+ \cos 2\Delta \psi + F_\times \sin 2\Delta\psi \, ,
\eeq
\beq
F_\times \rightarrow F_\times' = F_\times \cos 2\Delta \psi - F_+ \sin 2\Delta\psi \, ,
\eeq
ensuring that the observable $h(t)$ is independent of the arbitrary angle $\psi$.

\subsection{Templated analyses}

In modeling waveforms, it is useful to tie the polarization frame to the geometry of the source.
For a compact binary, it is conventional to compute the polarizations in a frame such that $\hat{x}$ lies along the intersection of the orbital plane with the plane of the sky (the line of nodes).
If the spins of the objects are a aligned and the source does not precess, one can use the Einstein field equation to show that, in that frame, $h_+$ and $h_\times$ take the form
\beq
h_+ = \left(1 + \cos^2 \iota\right) A(t; R, \vec{\theta}) \cos \Phi(t; \vec{\theta}) \, , 
\eeq
\beq
h_+ = 2 \cos \iota\, A(t; R, \vec{\theta}) \sin \Phi(t; \vec{\theta})\, ,
\eeq
where $A(t; \vec{\theta}) \equiv R  A(t; R, \vec{\theta})$ and $\Phi(t; \vec{\theta})$ are amplitude and phasing functions fully determined by intrinsic source parameters $\vec{\theta}$, which include the component masses, spin vectors.
The inclination $\iota$ is the polar angle with respect to the orbital angular momentum.
Importantly, for nonprecessing sources the plus and cross polarizations identical up to the inclination factors and a $\pi/2$ phase shift.

Having specified $h_+$ and $h_\times$ in this frame determines the values in any other frame via \eq{htransf}.
Since the waveform model will be prescribed in the specific frame above, we have to measure the polarization angle $\psi$, and we can think of it as a property of the source, rather than an arbitrary frame.
However, it is useful to keep in mind that this is a conceptual shortcut.

\subsection{Untemplated analyses}

If we do not have a model for the source, we can simply parametrize.

\section{Nontensor polarizations}

\newcommand{\hx}{h_{\rm x}}
\newcommand{\hy}{h_{\rm y}}
\newcommand{\hb}{h_{\rm b}}
\newcommand{\hlon}{h_{\rm l}}

In GR, there exist two propagating degrees of freedom, corresponding to two independent GW polarizations.
Their local effect can be encoded in a spatial driving matrix $h_{ij}$ that, in a frame with $z$-axis along the direction of propagation, we can write as
\beq \label{eq:polarizations}
(h_{ij}) = \begin{pmatrix}
\hb + h_+ & h_\times  & \hx  \\
h_\times  & \hb - h_+ & \hy  \\
\hx    & \hy    & \hlon
\end{pmatrix} ,
\eeq
with $h_p$, for $p$ in $\{+,\, \times,\, {\rm x},\, {\rm y},\, {\rm b},\, {\rm l}\}$, unspecified functions of time representing the six possible polarization degrees of freedom.
In GR, only the two tensor polarizations, plus ($+$) and cross ($\times$), are allowed, so that $\hx=\hy=\hb=\hlon=0$.
Different theories may also allow vector modes, x and y (also known as $v_1$ and $v_2$), or scalar modes, breathing (b) and longitudinal (l).

Equation \eqref{eq:polarizations} presumes a specific choice of frame orientation.
For $h_{ij}$ to take this simple form, it is necessary for the $z$ axis to be parallel to the (spatial) wave vector $\vec{k}$.
However, there is no restriction regarding the orientation of the $x$ and $y$ axes within the plane perpendicular to $\vec{k}$; this is usually specified by an arbitrary \emph{polarization angle} $\psi$, defined with respect to some convenient reference (e.g., the celestial equator).
The tensor, vector and scalar components of $h_{ij}$ are constructed to be invariant under $\psi$ rotations, with spin-weights $s=2$, $s=1$ and $s=0$ respectively.
This means that, after redefinitions
\beq
h_+ \rightarrow h_+' = h_+ \cos 2\Delta \psi - h_\times \sin 2\Delta\psi 
\eeq
\beq
h_\times \rightarrow h_\times' = h_\times \cos 2\Delta \psi + h_+ \sin 2\Delta\psi 
\eeq
\beq
\hx \rightarrow \hy' = \hx \cos \Delta \psi - \hy \sin \Delta\psi 
\eeq
\beq
\hy \rightarrow \hx' = \hy \cos \Delta \psi + \hx \sin \Delta\psi 
\eeq

\section{Conclusion}

\begin{acknowledgments}
% NASA
M.I.\ is supported by NASA through the NASA Hubble Fellowship
grant No.\ HST-HF2-51410.001-A awarded by the Space Telescope
Science Institute, which is operated by the Association of Universities
for Research in Astronomy, Inc., for NASA, under contract NAS5-26555.
% % LIGO
% M.I.\ is a member of the LIGO Laboratory.
% LIGO was constructed by the California Institute of Technology and
% Massachusetts Institute of Technology with funding from the National
% Science Foundation and operates under cooperative agreement PHY-0757058.
% DCC
This paper carries LIGO document number \dcc{}.
\end{acknowledgments}

\bibliography{refs}

\end{document}
%
% ****** End of file apstemplate.tex ******
